### 2025-04-17 analysis/supplementary-materials/appendix1.qmd
SOM SEM graphs finished in R ..02 Starting to add them to final SOM figs in illustrator.

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/SOM"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style

---

# Supplementary materials 1

## XRD results

In this subsection we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC. The original data necessary to produce these visualisations can be found and freely accessed in our online compendium, in the associated GitHub project under the path "../data/xrd_data".

```{r}
#| label: somxrd-library

library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(ggpmisc)

```

```{r}
#| label: somxrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

xrd_marks <- xrd_marks |> # Correct the quartz denomination on base dataset
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# Read individual xrd files
sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
sp18_xrd <- read_csv('data/xrd_data/SP18_xrd.csv')
sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')

```

```{r}
#| label: fig-xrd-sp6
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP6_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

# SP6 Cabo S. Vicente
sp6_marks <- xrd_marks |>
  filter(Sample == "SP6")
sp6_perc <- xrd_perc |>
  filter(Sample == "SP6") |>
  select(-Sample, -ESD)

sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))

```

```{r}
#| label: fig-xrd-sp7
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP7_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp7_marks <- xrd_marks |>
  filter(Sample == "SP7")
sp7_perc <- xrd_perc |>
  filter(Sample == "SP7") |>
  select(-Sample, -ESD)

sp7_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = `Quartz (SiO2)`) |> 
  rename(Dolomite = `Dolomite (CaMg(CO3)2)`) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp7_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp7_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp7_perc)))

```

```{r}
#| label: fig-xrd-sp10
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP10_PdA). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp10_marks <- xrd_marks |>
  filter(Sample == "SP10")
sp10_perc <- xrd_perc |>
  filter(Sample == "SP10") |>
  select(-Sample, -ESD)

sp10_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp10_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp10_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp10_perc)))

```

```{r}
#| label: fig-xrd-sp14
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP14_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp14_marks <- xrd_marks |>
  filter(Sample == "SP14")
sp14_perc <- xrd_perc |>
  filter(Sample == "SP14") |>
  select(-Sample, -ESD)

sp14_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp14_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp14_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp14_perc)))

```

```{r}
#| label: fig-xrd-sp18
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp18_marks <- xrd_marks |>
  filter(Sample == "SP18")
sp18_perc <- xrd_perc |>
  filter(Sample == "SP18") |>
  select(-Sample, -ESD)

sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))

```

```{r}
#| label: fig-xrd-sp21
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Belixe Sul, north of Praia do Belixe (SP21_BLS). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp21_marks <- xrd_marks |>
  filter(Sample == "SP21")
sp21_perc <- xrd_perc |>
  filter(Sample == "SP21") |>
  select(-Sample, -ESD)

sp21_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp21_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp21_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp21_perc)))

```

```{r}
#| label: fig-xrd-sp28
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Aspa, north of Cabo S. Vicente (SP28_ASP). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp28_marks <- xrd_marks |>
  filter(Sample == "SP28")
sp28_perc <- xrd_perc |>
  filter(Sample == "SP28") |>
  select(-Sample, -ESD)

sp28_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp28_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp28_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp28_perc)))

```

```{r}
#| label: fig-xrd-sp30
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP30_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp30_marks <- xrd_marks |>
  filter(Sample == "SP30")
sp30_perc <- xrd_perc |>
  filter(Sample == "SP30") |>
  select(-Sample, -ESD)

sp30_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp30_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp30_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp30_perc)))

```

```{r}
#| label: fig-xrd-sp33
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP33_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp33_marks <- xrd_marks |>
  filter(Sample == "SP33")
sp33_perc <- xrd_perc |>
  filter(Sample == "SP33") |>
  select(-Sample, -ESD)

sp33_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp33_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp33_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp33_perc)))

```

```{r}
#| label: fig-xrd-sp34
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP34_PdA_a). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp34_marks <- xrd_marks |>
  filter(Sample == "SP34")
sp34_perc <- xrd_perc |>
  filter(Sample == "SP34") |>
  select(-Sample, -ESD)

sp34_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp34_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp34_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp34_perc)))

```

```{r}
#| label: fig-xrd-sp40
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ferrel (SP40_b). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp40_marks <- xrd_marks |>
  filter(Sample == "SP40")
sp40_perc <- xrd_perc |>
  filter(Sample == "SP40") |>
  select(-Sample, -ESD)

sp40_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = SiO2) |> 
  select(-SiO2_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp40_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp40_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp40_perc)))

```

```{r}
#| label: fig-xrd-sp50
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP50_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp50_marks <- xrd_marks |>
  filter(Sample == "SP50")
sp50_perc <- xrd_perc |>
  filter(Sample == "SP50") |>
  select(-Sample, -ESD)

sp50_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp50_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp50_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp50_perc)))

```

```{r}
#| label: fig-xrd-sp52
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP52_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp52_marks <- xrd_marks |>
  filter(Sample == "SP52")
sp52_perc <- xrd_perc |>
  filter(Sample == "SP52") |>
  select(-Sample, -ESD)

sp52_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite ,Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp52_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp52_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp52_perc)))

```

```{r}
#| label: fig-xrd-sp57
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formations) chert sample from the outcrop Jordana (SP57_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp57_marks <- xrd_marks |>
  filter(Sample == "SP57")
sp57_perc <- xrd_perc |>
  filter(Sample == "SP57") |>
  select(-Sample, -ESD)

sp57_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp57_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp57_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp57_perc)))

```

```{r}
#| label: fig-xrd-sp59
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP59_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp59_marks <- xrd_marks |>
  filter(Sample == "SP59")
sp59_perc <- xrd_perc |>
  filter(Sample == "SP59") |>
  select(-Sample, -ESD)

sp59_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp59_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp59_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp59_perc)))

```

```{r}
#| label: fig-xrd-sp61
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the secondary outcrop Peral (SP61_PER). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp61_marks <- xrd_marks |>
  filter(Sample == "SP61")
sp61_perc <- xrd_perc |>
  filter(Sample == "SP61") |>
  select(-Sample, -ESD)

sp61_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp61_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp61_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp61_perc)))

```

```{r}
#| label: fig-xrd-sp63
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Oliveiras (SP63_OLV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp63_marks <- xrd_marks |>
  filter(Sample == "SP63")
sp63_perc <- xrd_perc |>
  filter(Sample == "SP63") |>
  select(-Sample, -ESD)

sp63_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp63_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp63_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp63_perc)))

```

```{r}
#| label: fig-xrd-sp66
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Malhão (SP66_MALH). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp66_marks <- xrd_marks |>
  filter(Sample == "SP66")
sp66_perc <- xrd_perc |>
  filter(Sample == "SP66") |>
  select(-Sample, -ESD)

sp66_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp66_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp66_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp66_perc)))

```

## SEM-EDS results

```{r}
#| label: sem-read-som

M2_3_map2 <- read_csv('analysis/data/raw_data/sem_data/M2_3_map2.csv')
M3_1 <- read_csv('analysis/data/raw_data/sem_data/M3_1_map1_map2.csv')
M3_3_s1 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s1.csv')
M3_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s2.csv')
M4_2_s5_1 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_1.csv')
M4_2_s5_4 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_4.csv')
M8_2 <- read_csv('analysis/data/raw_data/sem_data/M8_2.csv')
M10_4 <- read_csv('analysis/data/raw_data/sem_data/M10_4.csv')
M11_1 <- read_csv('analysis/data/raw_data/sem_data/M11_1.csv')
M12_1_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s2_1.csv')
M12_1_s3_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s3_1.csv')
M12_3_s3_2 <- read_csv('analysis/data/raw_data/sem_data/M12_3_s3_2.csv')
M12_5_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_1.csv')
M12_5_s1_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_3.csv')
M12_5_s3_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s3_3.csv')
M12_5_s6_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_1.csv')
M12_5_s6_2 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_2.csv')
M13_1_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M13_1_3_s2.csv')
M14_3_2_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s1_1.csv')
M14_3_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s2_1.csv')
M14_4_1 <- read_csv('analysis/data/raw_data/sem_data/M14_4_1.csv')
M15_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_1.csv')
M15_2_s2_2 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_2.csv')
M16_1_1 <- read_csv('analysis/data/raw_data/sem_data/M16_1_1.csv')
M18_1_1 <- read_csv('analysis/data/raw_data/sem_data/M18_1_1.csv')
M18_5_1 <- read_csv('analysis/data/raw_data/sem_data/M18_5_1.csv')
M20_2_1 <- read_csv('analysis/data/raw_data/sem_data/M20_2_1.csv')
M22_4.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_4.1_1.csv')
M22_5.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_1.csv')
M22_5.1_3 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_3.csv')
M23_2_1 <- read_csv('analysis/data/raw_data/sem_data/M23_2_1.csv')
M23_7_1 <- read_csv('analysis/data/raw_data/sem_data/M23_7_1.csv')
M24_1_1 <- read_csv('analysis/data/raw_data/sem_data/M24_1_1.csv')
M24_1_2 <- read_csv('analysis/data/raw_data/sem_data/M24_1_2.csv')

```

```{r}
#| label: sem-graphs-som

# Prepare the energy ranges
energy_ranges <- list(
  C = c(0.2, 0.3), # "#584B9F" dark blue
  O = c(0.5, 0.6), # "#0086B3" blue
  F = c(0.67, 0.69), # "#3EADA9"  greenish blue
  Na = c(1.04, 1.06), # "#4FC46A"  green"
  Mg = c(1.25, 1.27), # "#AEC491" yellowish green
  Mg_b = c(1.3, 1.4), # "#AEC491" yellowish green
  Al = c(1.48, 1.5), # "#F2C86C" yellow
  Si = c(1.7, 1.8), # "#F5A838" orange
  P = c(2, 2.1), # "#EE850E" dark orange
  S = c(2.3, 2.32), # "#D54F34" red
  S_b = c(2.46, 2.47), # "#D54F34" red
  Cl = c(2.62, 2.64), # "#A51122" wine
  K = c(3.31, 3.33), # "#A71B4B" pinkish red
  K_b = c(3.59, 3.61), # "#A71B4B" pinkish red
  Ca = c(3.68, 3.7), # "#D35C79" pink
  Ca_b = c(4.01, 4.02), # "#D35C79" pink
  Ca_L = c(0.34, 0.36), # "#D35C79" pink
  Ti = c(4.5, 4.6), # "#BD3977" light purple
  Ti_b = c(4.9, 5), # "#BD3977" light purple
  Ti_L = c(0.45, 0.47), # "#BD3977" light purple
  Fe = c(6.38, 6.41), # "#762181" purple
  Fe_b = c(7, 7.2), # "#762181" purple
  Fe_L = c(0.69, 0.73)) # "#762181" purple

# M24_1_2
filtered_M24_1_2 <- subset(M24_1_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M24_1_2 <- list()
for (element in names(energy_ranges)) {
  range_M24_1_2 <- energy_ranges[[element]]
  subset_M24_1_2 <- subset(filtered_M24_1_2, Energy >= range_M24_1_2[1] & Energy <= range_M24_1_2[2])
  if (nrow(subset_M24_1_2) > 0) {
    max_row_M24_1_2 <- subset_M24_1_2[which.max(subset_M24_1_2$Counts), ]
    peaks_list_M24_1_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M24_1_2$Energy,
      Counts = max_row_M24_1_2$Counts
    )
  }
}

peaks_M24_1_2 <- do.call(rbind, peaks_list_M24_1_2)
rownames(peaks_M24_1_2) <- NULL
peaks_M24_1_2 <- peaks_M24_1_2 |> 
  filter(Counts > 250) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S_b", "S", "Cl"))
peaks_M24_1_2_labels <- peaks_M24_1_2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b"))

M24_1_2_graph <- filtered_M24_1_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 9500)) +
  geom_point(data = peaks_M24_1_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M24_1_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M24_1_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M24_1_1
filtered_M24_1_1 <- subset(M24_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M24_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M24_1_1 <- energy_ranges[[element]]
  subset_M24_1_1 <- subset(filtered_M24_1_1, Energy >= range_M24_1_1[1] & Energy <= range_M24_1_1[2])
  if (nrow(subset_M24_1_1) > 0) {
    max_row_M24_1_1 <- subset_M24_1_1[which.max(subset_M24_1_1$Counts), ]
    peaks_list_M24_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M24_1_1$Energy,
      Counts = max_row_M24_1_1$Counts
    )
  }
}

peaks_M24_1_1 <- do.call(rbind, peaks_list_M24_1_1)
rownames(peaks_M24_1_1) <- NULL
peaks_M24_1_1 <- peaks_M24_1_1 |> 
  filter(Counts > 250) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P", "S", "Cl", "S_b", "K", "K_b", "Mg"))
peaks_M24_1_1_labels <- peaks_M24_1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b"))

M24_1_1_graph <- filtered_M24_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 8500)) +
  geom_point(data = peaks_M24_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M24_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M24_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M23_7_1
filtered_M23_7_1 <- subset(M23_7_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M23_7_1 <- list()
for (element in names(energy_ranges)) {
  range_M23_7_1 <- energy_ranges[[element]]
  subset_M23_7_1 <- subset(filtered_M23_7_1, Energy >= range_M23_7_1[1] & Energy <= range_M23_7_1[2])
  if (nrow(subset_M23_7_1) > 0) {
    max_row_M23_7_1 <- subset_M23_7_1[which.max(subset_M23_7_1$Counts), ]
    peaks_list_M23_7_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M23_7_1$Energy,
      Counts = max_row_M23_7_1$Counts
    )
  }
}

peaks_M23_7_1 <- do.call(rbind, peaks_list_M23_7_1)
rownames(peaks_M23_7_1) <- NULL
peaks_M23_7_1 <- peaks_M23_7_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M23_7_1_labels <- peaks_M23_7_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M23_7_1_graph <- filtered_M23_7_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 2500)) +
  geom_point(data = peaks_M23_7_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M23_7_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M23_7_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M23_2_1
filtered_M23_2_1 <- subset(M23_2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M23_2_1 <- list()
for (element in names(energy_ranges)) {
  range_M23_2_1 <- energy_ranges[[element]]
  subset_M23_2_1 <- subset(filtered_M23_2_1, Energy >= range_M23_2_1[1] & Energy <= range_M23_2_1[2])
  if (nrow(subset_M23_2_1) > 0) {
    max_row_M23_2_1 <- subset_M23_2_1[which.max(subset_M23_2_1$Counts), ]
    peaks_list_M23_2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M23_2_1$Energy,
      Counts = max_row_M23_2_1$Counts
    )
  }
}

peaks_M23_2_1 <- do.call(rbind, peaks_list_M23_2_1)
rownames(peaks_M23_2_1) <- NULL
peaks_M23_2_1 <- peaks_M23_2_1 |> 
  filter(Counts > 70) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
peaks_M23_2_1_labels <- peaks_M23_2_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M23_2_1_graph <- filtered_M23_2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 2200)) +
  geom_point(data = peaks_M23_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M23_2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M23_2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M22_5.1_3
filtered_M22_5.1_3 <- subset(M22_5.1_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_5.1_3 <- list()
for (element in names(energy_ranges)) {
  range_M22_5.1_3 <- energy_ranges[[element]]
  subset_M22_5.1_3 <- subset(filtered_M22_5.1_3, Energy >= range_M22_5.1_3[1] & Energy <= range_M22_5.1_3[2])
  if (nrow(subset_M22_5.1_3) > 0) {
    max_row_M22_5.1_3 <- subset_M22_5.1_3[which.max(subset_M22_5.1_3$Counts), ]
    peaks_list_M22_5.1_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_5.1_3$Energy,
      Counts = max_row_M22_5.1_3$Counts
    )
  }
}

peaks_M22_5.1_3 <- do.call(rbind, peaks_list_M22_5.1_3)
rownames(peaks_M22_5.1_3) <- NULL
peaks_M22_5.1_3 <- peaks_M22_5.1_3 |> 
  filter(Counts > 55) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "Cl"))
peaks_M22_5.1_3_labels <- peaks_M22_5.1_3 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M22_5.1_3_graph <- filtered_M22_5.1_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 1800)) +
  geom_point(data = peaks_M22_5.1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M22_5.1_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M22_5.1_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M22_5.1_1
filtered_M22_5.1_1 <- subset(M22_5.1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_5.1_1 <- list()
for (element in names(energy_ranges)) {
  range_M22_5.1_1 <- energy_ranges[[element]]
  subset_M22_5.1_1 <- subset(filtered_M22_5.1_1, Energy >= range_M22_5.1_1[1] & Energy <= range_M22_5.1_1[2])
  if (nrow(subset_M22_5.1_1) > 0) {
    max_row_M22_5.1_1 <- subset_M22_5.1_1[which.max(subset_M22_5.1_1$Counts), ]
    peaks_list_M22_5.1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_5.1_1$Energy,
      Counts = max_row_M22_5.1_1$Counts
    )
  }
}

peaks_M22_5.1_1 <- do.call(rbind, peaks_list_M22_5.1_1)
rownames(peaks_M22_5.1_1) <- NULL
peaks_M22_5.1_1 <- peaks_M22_5.1_1 |> 
  filter(Counts > 70) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S"))
peaks_M22_5.1_1_labels <- peaks_M22_5.1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M22_5.1_1_graph <- filtered_M22_5.1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 3800)) +
  geom_point(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M22_5.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M22_5.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M22_4.1_1
filtered_M22_4.1_1 <- subset(M22_4.1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_4.1_1 <- list()
for (element in names(energy_ranges)) {
  range_M22_4.1_1 <- energy_ranges[[element]]
  subset_M22_4.1_1 <- subset(filtered_M22_4.1_1, Energy >= range_M22_4.1_1[1] & Energy <= range_M22_4.1_1[2])
  if (nrow(subset_M22_4.1_1) > 0) {
    max_row_M22_4.1_1 <- subset_M22_4.1_1[which.max(subset_M22_4.1_1$Counts), ]
    peaks_list_M22_4.1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_4.1_1$Energy,
      Counts = max_row_M22_4.1_1$Counts
    )
  }
}

peaks_M22_4.1_1 <- do.call(rbind, peaks_list_M22_4.1_1)
rownames(peaks_M22_4.1_1) <- NULL
peaks_M22_4.1_1 <- peaks_M22_4.1_1 |> 
  filter(Counts > 50) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M22_4.1_1_labels <- peaks_M22_4.1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M22_4.1_1_graph <- filtered_M22_4.1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M22_4.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M22_4.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M20_2_1
filtered_M20_2_1 <- subset(M20_2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M20_2_1 <- list()
for (element in names(energy_ranges)) {
  range_M20_2_1 <- energy_ranges[[element]]
  subset_M20_2_1 <- subset(filtered_M20_2_1, Energy >= range_M20_2_1[1] & Energy <= range_M20_2_1[2])
  if (nrow(subset_M20_2_1) > 0) {
    max_row_M20_2_1 <- subset_M20_2_1[which.max(subset_M20_2_1$Counts), ]
    peaks_list_M20_2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M20_2_1$Energy,
      Counts = max_row_M20_2_1$Counts
    )
  }
}

peaks_M20_2_1 <- do.call(rbind, peaks_list_M20_2_1)
rownames(peaks_M20_2_1) <- NULL
peaks_M20_2_1 <- peaks_M20_2_1 |> 
  filter(Counts > 400) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S", "Mg", "S_b", "Ca_b", "Ti_b", "K_b", "K"))
peaks_M20_2_1_labels <- peaks_M20_2_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M20_2_1_graph <- filtered_M20_2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M20_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M20_2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M20_2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M18_5_1
filtered_M18_5_1 <- subset(M18_5_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_5_1 <- list()
for (element in names(energy_ranges)) {
  range_M18_5_1 <- energy_ranges[[element]]
  subset_M18_5_1 <- subset(filtered_M18_5_1, Energy >= range_M18_5_1[1] & Energy <= range_M18_5_1[2])
  if (nrow(subset_M18_5_1) > 0) {
    max_row_M18_5_1 <- subset_M18_5_1[which.max(subset_M18_5_1$Counts), ]
    peaks_list_M18_5_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_5_1$Energy,
      Counts = max_row_M18_5_1$Counts
    )
  }
}

peaks_M18_5_1 <- do.call(rbind, peaks_list_M18_5_1)
rownames(peaks_M18_5_1) <- NULL
peaks_M18_5_1 <- peaks_M18_5_1 |> 
  filter(Counts > 150) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M18_5_1_labels <- peaks_M18_5_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M18_5_1_graph <- filtered_M18_5_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M18_5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M18_5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M18_1_1
filtered_M18_1_1 <- subset(M18_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M18_1_1 <- energy_ranges[[element]]
  subset_M18_1_1 <- subset(filtered_M18_1_1, Energy >= range_M18_1_1[1] & Energy <= range_M18_1_1[2])
  if (nrow(subset_M18_1_1) > 0) {
    max_row_M18_1_1 <- subset_M18_1_1[which.max(subset_M18_1_1$Counts), ]
    peaks_list_M18_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_1_1$Energy,
      Counts = max_row_M18_1_1$Counts
    )
  }
}

peaks_M18_1_1 <- do.call(rbind, peaks_list_M18_1_1)
rownames(peaks_M18_1_1) <- NULL
peaks_M18_1_1 <- peaks_M18_1_1 |> 
  filter(Counts > 200) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M18_1_1_labels <- peaks_M18_1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M18_1_1_graph <- filtered_M18_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M18_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M18_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M16_1_1
filtered_M16_1_1 <- subset(M16_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M16_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M16_1_1 <- energy_ranges[[element]]
  subset_M16_1_1 <- subset(filtered_M16_1_1, Energy >= range_M16_1_1[1] & Energy <= range_M16_1_1[2])
  if (nrow(subset_M16_1_1) > 0) {
    max_row_M16_1_1 <- subset_M16_1_1[which.max(subset_M16_1_1$Counts), ]
    peaks_list_M16_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M16_1_1$Energy,
      Counts = max_row_M16_1_1$Counts
    )
  }
}

peaks_M16_1_1 <- do.call(rbind, peaks_list_M16_1_1)
rownames(peaks_M16_1_1) <- NULL
peaks_M16_1_1 <- peaks_M16_1_1 |> 
  filter(Counts > 600) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S_b"))
peaks_M16_1_1_labels <- peaks_M16_1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M16_1_1_graph <- filtered_M16_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M16_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M16_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M16_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M15_2_s2_2
filtered_M15_2_s2_2 <- subset(M15_2_s2_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M15_2_s2_2 <- list()
for (element in names(energy_ranges)) {
  range_M15_2_s2_2 <- energy_ranges[[element]]
  subset_M15_2_s2_2 <- subset(filtered_M15_2_s2_2, Energy >= range_M15_2_s2_2[1] & Energy <= range_M15_2_s2_2[2])
  if (nrow(subset_M15_2_s2_2) > 0) {
    max_row_M15_2_s2_2 <- subset_M15_2_s2_2[which.max(subset_M15_2_s2_2$Counts), ]
    peaks_list_M15_2_s2_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M15_2_s2_2$Energy,
      Counts = max_row_M15_2_s2_2$Counts
    )
  }
}

peaks_M15_2_s2_2 <- do.call(rbind, peaks_list_M15_2_s2_2)
rownames(peaks_M15_2_s2_2) <- NULL
peaks_M15_2_s2_2 <- peaks_M15_2_s2_2 |> 
  filter(Counts > 300) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "Na", "P", "S", "Cl", "K", "Ca_b", "S_b", "Ca", "K_b", "Ca_L", "Mg"))
peaks_M15_2_s2_2_labels <- peaks_M15_2_s2_2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M15_2_s2_2_graph <- filtered_M15_2_s2_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M15_2_s2_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M15_2_s2_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M15_2_s2_1
filtered_M15_2_s2_1 <- subset(M15_2_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M15_2_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M15_2_s2_1 <- energy_ranges[[element]]
  subset_M15_2_s2_1 <- subset(filtered_M15_2_s2_1, Energy >= range_M15_2_s2_1[1] & Energy <= range_M15_2_s2_1[2])
  if (nrow(subset_M15_2_s2_1) > 0) {
    max_row_M15_2_s2_1 <- subset_M15_2_s2_1[which.max(subset_M15_2_s2_1$Counts), ]
    peaks_list_M15_2_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M15_2_s2_1$Energy,
      Counts = max_row_M15_2_s2_1$Counts
    )
  }
}

peaks_M15_2_s2_1 <- do.call(rbind, peaks_list_M15_2_s2_1)
rownames(peaks_M15_2_s2_1) <- NULL
peaks_M15_2_s2_1 <- peaks_M15_2_s2_1 |> 
  filter(Counts > 500) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b", "Na", "Mg"))
peaks_M15_2_s2_1_labels <- peaks_M15_2_s2_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M15_2_s2_1_graph <- filtered_M15_2_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M15_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M15_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M14_4_1
filtered_M14_4_1 <- subset(M14_4_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_4_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_4_1 <- energy_ranges[[element]]
  subset_M14_4_1 <- subset(filtered_M14_4_1, Energy >= range_M14_4_1[1] & Energy <= range_M14_4_1[2])
  if (nrow(subset_M14_4_1) > 0) {
    max_row_M14_4_1 <- subset_M14_4_1[which.max(subset_M14_4_1$Counts), ]
    peaks_list_M14_4_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_4_1$Energy,
      Counts = max_row_M14_4_1$Counts
    )
  }
}

peaks_M14_4_1 <- do.call(rbind, peaks_list_M14_4_1)
rownames(peaks_M14_4_1) <- NULL
peaks_M14_4_1 <- peaks_M14_4_1 |> 
  filter(Counts > 700) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P"))
peaks_M14_4_1_labels <- peaks_M14_4_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M14_4_1_graph <- filtered_M14_4_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_4_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M14_4_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M14_4_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M14_3_2_s2_1
filtered_M14_3_2_s2_1 <- subset(M14_3_2_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_3_2_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_3_2_s2_1 <- energy_ranges[[element]]
  subset_M14_3_2_s2_1 <- subset(filtered_M14_3_2_s2_1, Energy >= range_M14_3_2_s2_1[1] & Energy <= range_M14_3_2_s2_1[2])
  if (nrow(subset_M14_3_2_s2_1) > 0) {
    max_row_M14_3_2_s2_1 <- subset_M14_3_2_s2_1[which.max(subset_M14_3_2_s2_1$Counts), ]
    peaks_list_M14_3_2_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_3_2_s2_1$Energy,
      Counts = max_row_M14_3_2_s2_1$Counts
    )
  }
}

peaks_M14_3_2_s2_1 <- do.call(rbind, peaks_list_M14_3_2_s2_1)
rownames(peaks_M14_3_2_s2_1) <- NULL
peaks_M14_3_2_s2_1 <- peaks_M14_3_2_s2_1 |> 
  filter(Counts > 60) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M14_3_2_s2_1_labels <- peaks_M14_3_2_s2_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M14_3_2_s2_1_graph <- filtered_M14_3_2_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M14_3_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M14_3_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M14_3_2_s1_1
filtered_M14_3_2_s1_1 <- subset(M14_3_2_s1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_3_2_s1_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_3_2_s1_1 <- energy_ranges[[element]]
  subset_M14_3_2_s1_1 <- subset(filtered_M14_3_2_s1_1, Energy >= range_M14_3_2_s1_1[1] & Energy <= range_M14_3_2_s1_1[2])
  if (nrow(subset_M14_3_2_s1_1) > 0) {
    max_row_M14_3_2_s1_1 <- subset_M14_3_2_s1_1[which.max(subset_M14_3_2_s1_1$Counts), ]
    peaks_list_M14_3_2_s1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_3_2_s1_1$Energy,
      Counts = max_row_M14_3_2_s1_1$Counts
    )
  }
}

peaks_M14_3_2_s1_1 <- do.call(rbind, peaks_list_M14_3_2_s1_1)
rownames(peaks_M14_3_2_s1_1) <- NULL
peaks_M14_3_2_s1_1 <- peaks_M14_3_2_s1_1 |> 
  filter(Counts > 500) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M14_3_2_s1_1_labels <- peaks_M14_3_2_s1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M14_3_2_s1_1_graph <- filtered_M14_3_2_s1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M14_3_2_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M14_3_2_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M13_1_3_s2
filtered_M13_1_3_s2 <- subset(M13_1_3_s2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M13_1_3_s2 <- list()
for (element in names(energy_ranges)) {
  range_M13_1_3_s2 <- energy_ranges[[element]]
  subset_M13_1_3_s2 <- subset(filtered_M13_1_3_s2, Energy >= range_M13_1_3_s2[1] & Energy <= range_M13_1_3_s2[2])
  if (nrow(subset_M13_1_3_s2) > 0) {
    max_row_M13_1_3_s2 <- subset_M13_1_3_s2[which.max(subset_M13_1_3_s2$Counts), ]
    peaks_list_M13_1_3_s2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M13_1_3_s2$Energy,
      Counts = max_row_M13_1_3_s2$Counts
    )
  }
}

peaks_M13_1_3_s2 <- do.call(rbind, peaks_list_M13_1_3_s2)
rownames(peaks_M13_1_3_s2) <- NULL
peaks_M13_1_3_s2 <- peaks_M13_1_3_s2 |> 
  filter(Counts > 300) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M13_1_3_s2_labels <- peaks_M13_1_3_s2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M13_1_3_s2_graph <- filtered_M13_1_3_s2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M13_1_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M13_1_3_s2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M13_1_3_s2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_5_s6_2
filtered_M12_5_s6_2 <- subset(M12_5_s6_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s6_2 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s6_2 <- energy_ranges[[element]]
  subset_M12_5_s6_2 <- subset(filtered_M12_5_s6_2, Energy >= range_M12_5_s6_2[1] & Energy <= range_M12_5_s6_2[2])
  if (nrow(subset_M12_5_s6_2) > 0) {
    max_row_M12_5_s6_2 <- subset_M12_5_s6_2[which.max(subset_M12_5_s6_2$Counts), ]
    peaks_list_M12_5_s6_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s6_2$Energy,
      Counts = max_row_M12_5_s6_2$Counts
    )
  }
}

peaks_M12_5_s6_2 <- do.call(rbind, peaks_list_M12_5_s6_2)
rownames(peaks_M12_5_s6_2) <- NULL
peaks_M12_5_s6_2 <- peaks_M12_5_s6_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_5_s6_2_labels <- peaks_M12_5_s6_2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_5_s6_2_graph <- filtered_M12_5_s6_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_5_s6_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_5_s6_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_5_s6_1
filtered_M12_5_s6_1 <- subset(M12_5_s6_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s6_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s6_1 <- energy_ranges[[element]]
  subset_M12_5_s6_1 <- subset(filtered_M12_5_s6_1, Energy >= range_M12_5_s6_1[1] & Energy <= range_M12_5_s6_1[2])
  if (nrow(subset_M12_5_s6_1) > 0) {
    max_row_M12_5_s6_1 <- subset_M12_5_s6_1[which.max(subset_M12_5_s6_1$Counts), ]
    peaks_list_M12_5_s6_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s6_1$Energy,
      Counts = max_row_M12_5_s6_1$Counts
    )
  }
}

peaks_M12_5_s6_1 <- do.call(rbind, peaks_list_M12_5_s6_1)
rownames(peaks_M12_5_s6_1) <- NULL
peaks_M12_5_s6_1 <- peaks_M12_5_s6_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_5_s6_1_labels <- peaks_M12_5_s6_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_5_s6_1_graph <- filtered_M12_5_s6_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_5_s6_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_5_s6_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_5_s3_3
filtered_M12_5_s3_3 <- subset(M12_5_s3_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s3_3 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s3_3 <- energy_ranges[[element]]
  subset_M12_5_s3_3 <- subset(filtered_M12_5_s3_3, Energy >= range_M12_5_s3_3[1] & Energy <= range_M12_5_s3_3[2])
  if (nrow(subset_M12_5_s3_3) > 0) {
    max_row_M12_5_s3_3 <- subset_M12_5_s3_3[which.max(subset_M12_5_s3_3$Counts), ]
    peaks_list_M12_5_s3_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s3_3$Energy,
      Counts = max_row_M12_5_s3_3$Counts
    )
  }
}

peaks_M12_5_s3_3 <- do.call(rbind, peaks_list_M12_5_s3_3)
rownames(peaks_M12_5_s3_3) <- NULL
peaks_M12_5_s3_3 <- peaks_M12_5_s3_3 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_5_s3_3_labels <- peaks_M12_5_s3_3 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_5_s3_3_graph <- filtered_M12_5_s3_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_5_s3_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_5_s3_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_5_s1_3
filtered_M12_5_s1_3 <- subset(M12_5_s1_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s1_3 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s1_3 <- energy_ranges[[element]]
  subset_M12_5_s1_3 <- subset(filtered_M12_5_s1_3, Energy >= range_M12_5_s1_3[1] & Energy <= range_M12_5_s1_3[2])
  if (nrow(subset_M12_5_s1_3) > 0) {
    max_row_M12_5_s1_3 <- subset_M12_5_s1_3[which.max(subset_M12_5_s1_3$Counts), ]
    peaks_list_M12_5_s1_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s1_3$Energy,
      Counts = max_row_M12_5_s1_3$Counts
    )
  }
}

peaks_M12_5_s1_3 <- do.call(rbind, peaks_list_M12_5_s1_3)
rownames(peaks_M12_5_s1_3) <- NULL
peaks_M12_5_s1_3 <- peaks_M12_5_s1_3 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_5_s1_3_labels <- peaks_M12_5_s1_3 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_5_s1_3_graph <- filtered_M12_5_s1_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_5_s1_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

#ggsave("M12_5_s1_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_5_s1_1
filtered_M12_5_s1_1 <- subset(M12_5_s1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s1_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s1_1 <- energy_ranges[[element]]
  subset_M12_5_s1_1 <- subset(filtered_M12_5_s1_1, Energy >= range_M12_5_s1_1[1] & Energy <= range_M12_5_s1_1[2])
  if (nrow(subset_M12_5_s1_1) > 0) {
    max_row_M12_5_s1_1 <- subset_M12_5_s1_1[which.max(subset_M12_5_s1_1$Counts), ]
    peaks_list_M12_5_s1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s1_1$Energy,
      Counts = max_row_M12_5_s1_1$Counts
    )
  }
}

peaks_M12_5_s1_1 <- do.call(rbind, peaks_list_M12_5_s1_1)
rownames(peaks_M12_5_s1_1) <- NULL
peaks_M12_5_s1_1 <- peaks_M12_5_s1_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L","Mg_b", "F"))
peaks_M12_5_s1_1_labels <- peaks_M12_5_s1_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_5_s1_1_graph <- filtered_M12_5_s1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_5_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_5_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_3_s3_2
filtered_M12_3_s3_2 <- subset(M12_3_s3_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_3_s3_2 <- list()
for (element in names(energy_ranges)) {
  range_M12_3_s3_2 <- energy_ranges[[element]]
  subset_M12_3_s3_2 <- subset(filtered_M12_3_s3_2, Energy >= range_M12_3_s3_2[1] & Energy <= range_M12_3_s3_2[2])
  if (nrow(subset_M12_3_s3_2) > 0) {
    max_row_M12_3_s3_2 <- subset_M12_3_s3_2[which.max(subset_M12_3_s3_2$Counts), ]
    peaks_list_M12_3_s3_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_3_s3_2$Energy,
      Counts = max_row_M12_3_s3_2$Counts
    )
  }
}

peaks_M12_3_s3_2 <- do.call(rbind, peaks_list_M12_3_s3_2)
rownames(peaks_M12_3_s3_2) <- NULL
peaks_M12_3_s3_2 <- peaks_M12_3_s3_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_3_s3_2_labels <- peaks_M12_3_s3_2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_3_s3_2_graph <- filtered_M12_3_s3_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_3_s3_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_3_s3_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_3_s3_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_1_s3_1
filtered_M12_1_s3_1 <- subset(M12_1_s3_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_1_s3_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_1_s3_1 <- energy_ranges[[element]]
  subset_M12_1_s3_1 <- subset(filtered_M12_1_s3_1, Energy >= range_M12_1_s3_1[1] & Energy <= range_M12_1_s3_1[2])
  if (nrow(subset_M12_1_s3_1) > 0) {
    max_row_M12_1_s3_1 <- subset_M12_1_s3_1[which.max(subset_M12_1_s3_1$Counts), ]
    peaks_list_M12_1_s3_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_1_s3_1$Energy,
      Counts = max_row_M12_1_s3_1$Counts
    )
  }
}

peaks_M12_1_s3_1 <- do.call(rbind, peaks_list_M12_1_s3_1)
rownames(peaks_M12_1_s3_1) <- NULL
peaks_M12_1_s3_1 <- peaks_M12_1_s3_1 |> 
  filter(Counts > 90) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_1_s3_1_labels <- peaks_M12_1_s3_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_1_s3_1_graph <- filtered_M12_1_s3_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_1_s3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_1_s3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M12_1_s2_1
filtered_M12_1_s2_1 <- subset(M12_1_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_1_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_1_s2_1 <- energy_ranges[[element]]
  subset_M12_1_s2_1 <- subset(filtered_M12_1_s2_1, Energy >= range_M12_1_s2_1[1] & Energy <= range_M12_1_s2_1[2])
  if (nrow(subset_M12_1_s2_1) > 0) {
    max_row_M12_1_s2_1 <- subset_M12_1_s2_1[which.max(subset_M12_1_s2_1$Counts), ]
    peaks_list_M12_1_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_1_s2_1$Energy,
      Counts = max_row_M12_1_s2_1$Counts
    )
  }
}

peaks_M12_1_s2_1 <- do.call(rbind, peaks_list_M12_1_s2_1)
rownames(peaks_M12_1_s2_1) <- NULL
peaks_M12_1_s2_1 <- peaks_M12_1_s2_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M12_1_s2_1_labels <- peaks_M12_1_s2_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M12_1_s2_1_graph <- filtered_M12_1_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M12_1_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M12_1_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M2_3_map2
filtered_M2_3_map2 <- subset(M2_3_map2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M2_3_map2 <- list()
for (element in names(energy_ranges)) {
  range_M2_3_map2 <- energy_ranges[[element]]
  subset_M2_3_map2 <- subset(filtered_M2_3_map2, Energy >= range_M2_3_map2[1] & Energy <= range_M2_3_map2[2])
  if (nrow(subset_M2_3_map2) > 0) {
    max_row_M2_3_map2 <- subset_M2_3_map2[which.max(subset_M2_3_map2$Counts), ]
    peaks_list_M2_3_map2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M2_3_map2$Energy,
      Counts = max_row_M2_3_map2$Counts
    )
  }
}

peaks_M2_3_map2 <- do.call(rbind, peaks_list_M2_3_map2)
rownames(peaks_M2_3_map2) <- NULL
peaks_M2_3_map2 <- peaks_M2_3_map2 |> 
  filter(Counts > 100) |> 
  filter(Element != "Ca_L")
peaks_M2_3_map2_labels <- peaks_M2_3_map2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M2_3_map2_graph <- filtered_M2_3_map2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M2_3_map2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M2_3_map2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M3_1
filtered_M3_1 <- subset(M3_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_1 <- list()
for (element in names(energy_ranges)) {
  range_M3_1 <- energy_ranges[[element]]
  subset_M3_1 <- subset(filtered_M3_1, Energy >= range_M3_1[1] & Energy <= range_M3_1[2])
  if (nrow(subset_M3_1) > 0) {
    max_row_M3_1 <- subset_M3_1[which.max(subset_M3_1$Counts), ]
    peaks_list_M3_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_1$Energy,
      Counts = max_row_M3_1$Counts
    )
  }
}

peaks_M3_1 <- do.call(rbind, peaks_list_M3_1)
rownames(peaks_M3_1) <- NULL
peaks_M3_1 <- peaks_M3_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "K_b"))
peaks_M3_1_labels <- peaks_M3_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M3_1_graph <- filtered_M3_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

ggsave("M3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M3_3_s1
filtered_M3_3_s1 <- subset(M3_3_s1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_3_s1 <- list()
for (element in names(energy_ranges)) {
  range_M3_3_s1 <- energy_ranges[[element]]
  subset_M3_3_s1 <- subset(filtered_M3_3_s1, Energy >= range_M3_3_s1[1] & Energy <= range_M3_3_s1[2])
  if (nrow(subset_M3_3_s1) > 0) {
    max_row_M3_3_s1 <- subset_M3_3_s1[which.max(subset_M3_3_s1$Counts), ]
    peaks_list_M3_3_s1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_3_s1$Energy,
      Counts = max_row_M3_3_s1$Counts
    )
  }
}

peaks_M3_3_s1 <- do.call(rbind, peaks_list_M3_3_s1)
rownames(peaks_M3_3_s1) <- NULL
peaks_M3_3_s1 <- peaks_M3_3_s1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "K_b", "Fe_L"))
peaks_M3_3_s1_labels <- peaks_M3_3_s1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M3_3_s1_graph <- filtered_M3_3_s1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 7.5)) +
  geom_point(data = peaks_M3_3_s1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M3_3_s1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M3_3_s1_graph.png", width = 10, height = 5, bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M3_3_s2
filtered_M3_3_s2 <- subset(M3_3_s2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_3_s2 <- list()
for (element in names(energy_ranges)) {
  range_M3_3_s2 <- energy_ranges[[element]]
  subset_M3_3_s2 <- subset(filtered_M3_3_s2, Energy >= range_M3_3_s2[1] & Energy <= range_M3_3_s2[2])
  if (nrow(subset_M3_3_s2) > 0) {
    max_row_M3_3_s2 <- subset_M3_3_s2[which.max(subset_M3_3_s2$Counts), ]
    peaks_list_M3_3_s2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_3_s2$Energy,
      Counts = max_row_M3_3_s2$Counts
    )
  }
}

peaks_M3_3_s2 <- do.call(rbind, peaks_list_M3_3_s2)
rownames(peaks_M3_3_s2) <- NULL
peaks_M3_3_s2 <- peaks_M3_3_s2 |> 
  filter(Counts > 90) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
peaks_M3_3_s2_labels <- peaks_M3_3_s2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M3_3_s2_graph <- filtered_M3_3_s2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M3_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M3_3_s2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M3_3_s2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M4_2_s5_1
filtered_M4_2_s5_1 <- subset(M4_2_s5_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_2_s5_1 <- list()
for (element in names(energy_ranges)) {
  range_M4_2_s5_1 <- energy_ranges[[element]]
  subset_M4_2_s5_1 <- subset(filtered_M4_2_s5_1, Energy >= range_M4_2_s5_1[1] & Energy <= range_M4_2_s5_1[2])
  if (nrow(subset_M4_2_s5_1) > 0) {
    max_row_M4_2_s5_1 <- subset_M4_2_s5_1[which.max(subset_M4_2_s5_1$Counts), ]
    peaks_list_M4_2_s5_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_2_s5_1$Energy,
      Counts = max_row_M4_2_s5_1$Counts
    )
  }
}

peaks_M4_2_s5_1 <- do.call(rbind, peaks_list_M4_2_s5_1)
rownames(peaks_M4_2_s5_1) <- NULL
peaks_M4_2_s5_1 <- peaks_M4_2_s5_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
peaks_M4_2_s5_1_labels <- peaks_M4_2_s5_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M4_2_s5_1_graph <- filtered_M4_2_s5_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(xlim = c(0.3, 5)) +
  geom_point(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M4_2_s5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

#ggsave("M4_2_s5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M4_2_s5_4
filtered_M4_2_s5_4 <- subset(M4_2_s5_4, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_2_s5_4 <- list()
for (element in names(energy_ranges)) {
  range_M4_2_s5_4 <- energy_ranges[[element]]
  subset_M4_2_s5_4 <- subset(filtered_M4_2_s5_4, Energy >= range_M4_2_s5_4[1] & Energy <= range_M4_2_s5_4[2])
  if (nrow(subset_M4_2_s5_4) > 0) {
    max_row_M4_2_s5_4 <- subset_M4_2_s5_4[which.max(subset_M4_2_s5_4$Counts), ]
    peaks_list_M4_2_s5_4[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_2_s5_4$Energy,
      Counts = max_row_M4_2_s5_4$Counts
    )
  }
}

peaks_M4_2_s5_4 <- do.call(rbind, peaks_list_M4_2_s5_4)
rownames(peaks_M4_2_s5_4) <- NULL
peaks_M4_2_s5_4 <- peaks_M4_2_s5_4 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M4_2_s5_4_labels <- peaks_M4_2_s5_4 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M4_2_s5_4_graph <- filtered_M4_2_s5_4 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M4_2_s5_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M4_2_s5_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M8_2
filtered_M8_2 <- subset(M8_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M8_2 <- list()
for (element in names(energy_ranges)) {
  range_M8_2 <- energy_ranges[[element]]
  subset_M8_2 <- subset(filtered_M8_2, Energy >= range_M8_2[1] & Energy <= range_M8_2[2])
  if (nrow(subset_M8_2) > 0) {
    max_row_M8_2 <- subset_M8_2[which.max(subset_M8_2$Counts), ]
    peaks_list_M8_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M8_2$Energy,
      Counts = max_row_M8_2$Counts
    )
  }
}

peaks_M8_2 <- do.call(rbind, peaks_list_M8_2)
rownames(peaks_M8_2) <- NULL
peaks_M8_2 <- peaks_M8_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
peaks_M8_2_labels <- peaks_M8_2 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M8_2_graph <- filtered_M8_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 5)) +
  geom_point(data = peaks_M8_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M8_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

 ggsave("M8_2_graph.png",width = 7, bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M10_4
filtered_M10_4 <- subset(M10_4, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M10_4 <- list()
for (element in names(energy_ranges)) {
  range_M10_4 <- energy_ranges[[element]]
  subset_M10_4 <- subset(filtered_M10_4, Energy >= range_M10_4[1] & Energy <= range_M10_4[2])
  if (nrow(subset_M10_4) > 0) {
    max_row_M10_4 <- subset_M10_4[which.max(subset_M10_4$Counts), ]
    peaks_list_M10_4[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M10_4$Energy,
      Counts = max_row_M10_4$Counts
    )
  }
}

peaks_M10_4 <- do.call(rbind, peaks_list_M10_4)
rownames(peaks_M10_4) <- NULL
peaks_M10_4 <- peaks_M10_4 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L","Mg_b", "F", "Ca_L", "Fe_L", "Na", "Mg", "P", "S", "Cl", "Ca", "Fe_b", "K", "K_b", "S_b"))
peaks_M10_4_labels <- peaks_M10_4 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M10_4_graph <- filtered_M10_4 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M10_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M10_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M10_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

# M11_1
filtered_M11_1 <- subset(M11_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M11_1 <- list()
for (element in names(energy_ranges)) {
  range_M11_1 <- energy_ranges[[element]]
  subset_M11_1 <- subset(filtered_M11_1, Energy >= range_M11_1[1] & Energy <= range_M11_1[2])
  if (nrow(subset_M11_1) > 0) {
    max_row_M11_1 <- subset_M11_1[which.max(subset_M11_1$Counts), ]
    peaks_list_M11_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M11_1$Energy,
      Counts = max_row_M11_1$Counts
    )
  }
}

peaks_M11_1 <- do.call(rbind, peaks_list_M11_1)
rownames(peaks_M11_1) <- NULL
peaks_M11_1 <- peaks_M11_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
peaks_M11_1_labels <- peaks_M11_1 |> 
  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))

M11_1_graph <- filtered_M11_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M11_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_text(data = peaks_M11_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))

# ggsave("M11_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

```

### Geological samples

#### Lower Jurassic of western Algarve

The analysed samples from the Lower Jurassic outcrops are homogeneous regarding their structure, inclusions and mineral compositions. The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of Western Algarve, including the inland outcrop of Ferrel).

They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and small peaks of aluminium, with a single sample also showing a high aluminium peak. Small peaks of chlorine and potassium can also be identified in less than 50% of the samples.

Phosphatised skeletal remains are very frequent, occurring in more than 80% of the samples. These are all composed of calcium and phosphorus, and frequently show an associated smaller peak of sulphur (\~50% of the samples) or sulphur and fluorine (\~70% of the samples). These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the phosphatised skeletal remains. These skeletal remains are frequently irregular and indeterminate; however, a small number of samples show phosphatised skeletal remains with homogeneous oval shapes, which are currently indeterminate.

Finally, iron oxides are present in all the samples, with an iron (Fe) composition. Other types of iron oxides are also present in lesser frequencies, as equally large peaks of iron and titanium (40% of the samples), iron and fluoride (20%), and titanium with a small peak of iron (\<10%). A single sample shows a high titanium peak without iron and another titanium with a smaller peak of phosphorus. Most of these iron or titanium inclusions have irregular shapes and varying sizes. Albeit rare (present in 3 samples), pyrite can be identified in the form of spherical framboids or dispersed pyrite nanocrystals.