### 2025-04-19 analysis/paper/paper.qmd
Small improvements to caps on main paper

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r}
#| label: libraries

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(ggfortify)
library(FactoMineR)
library(factoextra)
library(ggpmisc)
library(magick)

```

```{r}
#| label: xrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

# Read individual xrd files
sp15_xrd <- read_csv('data/xrd_data/SP15_xrd.csv')
sp58_xrd <- read_csv('data/xrd_data/SP58_xrd.csv')
sp62_xrd <- read_csv('data/xrd_data/SP62_xrd.csv')
sp69_xrd <- read_csv('data/xrd_data/SP69_xrd.csv')

```

```{r}
#| label: xrd-graphs

xrd_marks <- xrd_marks |>
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# Colour codes
## Observed = "lightsteelblue3"
## Quartz = "salmon"
## Dolomite = "steelblue1"
## Calcite = "darkseagreen"
## Halite = "darkgoldenrod1"
## Background = "orchid"
## Difference = "snow4"

# SP15
sp15_marks <- xrd_marks |>
  filter(Sample == "SP15")
sp15_perc <- xrd_perc |>
  filter(Sample == "SP15") |>
  select(-Sample, -ESD)

sp15_plot <- sp15_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp15_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp15_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp15_perc)))

# SP58 Jordana
sp58_marks <- xrd_marks |>
  filter(Sample == "SP58")
sp58_perc <- xrd_perc |>
  filter(Sample == "SP58") |>
  select(-Sample, -ESD)

sp58_plot <- sp58_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp58_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp58_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp58_perc)))

# SP62 Oliveiras
sp62_marks <- xrd_marks |> 
  filter(Sample == "SP62")
sp62_perc <- xrd_perc |> 
  filter(Sample == "SP62") |> 
  select(-Sample, -ESD)

sp62_plot <- sp62_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp62_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp62_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp62_perc)))

# SP69 Mareta
sp69_marks <- xrd_marks |> 
  filter(Sample == "SP69")
sp69_perc <- xrd_perc |> 
  filter(Sample == "SP69") |> 
  select(-Sample, -ESD)

sp69_plot <- sp69_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp69_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp69_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp69_perc)))

```

```{r}
#| label: pxrf-read

pxrf_raw <- read_csv('data/pxrf_data_dec.csv')

```

```{r}
#| label: sem-read

M18_7 <- read_csv('analysis/data/raw_data/sem_data/M18_7.csv')
M18_8 <- read_csv('analysis/data/raw_data/sem_data/M18_8.csv')
M4_3 <- read_csv('analysis/data/raw_data/sem_data/M4_3.csv')

```

```{r}
#| label: sem-graphs

# Prepare the energy ranges
energy_ranges <- list(
  C = c(0.2, 0.3), # "#584B9F" dark blue
  O = c(0.5, 0.6), # "#0086B3" blue
  F = c(0.67, 0.69), # "#3EADA9"  greenish blue
  Na = c(1.04, 1.06), # "#4FC46A"  green"
  Mg = c(1.25, 1.27), # "#AEC491" yellowish green
  Mg_b = c(1.3, 1.4), # "#AEC491" yellowish green
  Al = c(1.48, 1.5), # "#F2C86C" yellow
  Si = c(1.7, 1.8), # "#F5A838" orange
  P = c(2, 2.1), # "#EE850E" dark orange
  S = c(2.3, 2.32), # "#D54F34" red
  S_b = c(2.46, 2.47), # "#D54F34" red
  Cl = c(2.62, 2.64), # "#A51122" wine
  K = c(3.31, 3.33), # "#A71B4B" pinkish red
  K_b = c(3.59, 3.61), # "#A71B4B" pinkish red
  Ca = c(3.68, 3.7), # "#D35C79" pink
  Ca_b = c(4.01, 4.02), # "#D35C79" pink
  Ca_L = c(0.34, 0.36), # "#D35C79" pink
  Ti = c(4.5, 4.6), # "#BD3977" light purple
  Ti_b = c(4.9, 5), # "#BD3977" light purple
  Ti_L = c(0.45, 0.47), # "#BD3977" light purple
  Fe = c(6.38, 6.41), # "#762181" purple
  Fe_b = c(7, 7.2), # "#762181" purple
  Fe_L = c(0.69, 0.73)) # "#762181" purple

# M18_8
filtered_M18_8 <- subset(M18_8, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_8 <- list()
for (element in names(energy_ranges)) {
  range_M18_8 <- energy_ranges[[element]]
  subset_M18_8 <- subset(filtered_M18_8, Energy >= range_M18_8[1] & Energy <= range_M18_8[2])
  if (nrow(subset_M18_8) > 0) {
    max_row_M18_8 <- subset_M18_8[which.max(subset_M18_8$Counts), ]
    peaks_list_M18_8[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_8$Energy,
      Counts = max_row_M18_8$Counts
    )
  }
}

peaks_M18_8 <- do.call(rbind, peaks_list_M18_8)
rownames(peaks_M18_8) <- NULL
peaks_M18_8 <- peaks_M18_8 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
peaks_M18_8_labels <- peaks_M18_8 |> 
  filter(Element != "Fe_L")

M18_8_graph <- filtered_M18_8 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 5000)) +
  geom_point(data = peaks_M18_8, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  #geom_label_repel(data = peaks_M18_8_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, fill = "white", label.size = 0.2, direction = "y", nudge_y = 0.1, max.overlaps = Inf, show.legend = FALSE) +
  geom_text(data = peaks_M18_8_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))

# ggsave("M18_8_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")

# M18_7
filtered_M18_7 <- subset(M18_7, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_7 <- list()
for (element in names(energy_ranges)) {
  range_M18_7 <- energy_ranges[[element]]
  subset_M18_7 <- subset(filtered_M18_7, Energy >= range_M18_7[1] & Energy <= range_M18_7[2])
  if (nrow(subset_M18_7) > 0) {
    max_row_M18_7 <- subset_M18_7[which.max(subset_M18_7$Counts), ]
    peaks_list_M18_7[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_7$Energy,
      Counts = max_row_M18_7$Counts
    )
  }
}

peaks_M18_7 <- do.call(rbind, peaks_list_M18_7)
rownames(peaks_M18_7) <- NULL
peaks_M18_7 <- peaks_M18_7 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("F", "Ti_L"))
peaks_M18_7_labels <- peaks_M18_7 |> 
  filter(!Element %in% c("Fe_L", "Fe_b"))

M18_7_graph <- filtered_M18_7 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 4000)) +
  geom_point(data = peaks_M18_7, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  #geom_label_repel(data = peaks_M18_8, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, fill = "white", label.size = 0.2, direction = "y", nudge_y = 0.1, max.overlaps = Inf, show.legend = FALSE) +
  geom_text(data = peaks_M18_7_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))

# ggsave("M18_7_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")

# M4_3
filtered_M4_3 <- subset(M4_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_3 <- list()
for (element in names(energy_ranges)) {
  range_M4_3 <- energy_ranges[[element]]
  subset_M4_3 <- subset(filtered_M4_3, Energy >= range_M4_3[1] & Energy <= range_M4_3[2])
  if (nrow(subset_M4_3) > 0) {
    max_row_M4_3 <- subset_M4_3[which.max(subset_M4_3$Counts), ]
    peaks_list_M4_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_3$Energy,
      Counts = max_row_M4_3$Counts
    )
  }
}

peaks_M4_3 <- do.call(rbind, peaks_list_M4_3)
rownames(peaks_M4_3) <- NULL
peaks_M4_3 <- peaks_M4_3 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Mg_b", "K_b", "Ti_L"))
peaks_M4_3_labels <- peaks_M4_3 |> 
  filter(Element != "Ca_b")

M4_3_graph <- filtered_M4_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 5000), xlim = c(0.3, 4.5)) +
  geom_point(data = peaks_M4_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  #geom_text_repel(data = peaks_M4_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, direction = "y", nudge_y = 0.5, max.overlaps = Inf, force = 1, show.legend = FALSE) +
  geom_text(data = peaks_M4_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
  scale_colour_manual(values = c(
  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +  
  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))

ggsave("M4_3_graph.png", width = 5, height = 3.75, units = "in", ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")

```

```{r}
#| label: pxrf-database-setup

## DB setup
pxrf_db <- pxrf_raw |> 
  filter(!Sample_ID %in% c("I19-1751", "H21-2126")) # Removing 2 altered samples
pxrf_analysis <- pxrf_db |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr") |>  # Removed elements with 0 or <LOD values in all samples or in high percentages in all groups
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2",
    `Formation/Type` == "TYPE 3" ~ "T3",
    `Formation/Type` == "TYPE 4" ~ "T4",
    `Formation/Type` == "TYPE 5" ~ "T5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  ))

```

```{r}
#| label: pxrf-fig-lod
#| fig-cap: A barplot.

# Table to see the occurrence of 0 or <LOD values within each element and by rm type, to choose the elements in which to apply statistics.
# Combine the multiple columns into a single table
lod_tab <- pxrf_db |>
  select(-contains("Err")) |>
  select(-pXRF_ID, -Application, -Method, -ElapsedTime, -`Cal Check`, -Sample_ID) |>
  mutate(across(.cols = MgO:U,
                .fns = ~ case_when(
                  . == 0 ~"Success 0",
                  . > 0 ~ "Success",
                  TRUE ~ as.character(.))
                ))
# Reshape the data into a long format
lod_tab_long <- lod_tab |>
  pivot_longer(cols = MgO:U,  # Assuming these columns contain the 'Success' and 'LOD' values
               names_to = "Element",  # New column for the names of elements
               values_to = "Value")   # New column for the corresponding Success/LOD values

lod_tab_summary <- lod_tab_long |>
  group_by(Element, Value) |>
  summarise(Count = n(), .groups = "drop")  # Count the occurrences of Success and LOD for each Element

# Reshape the data into a wide format using pivot_wider
lod_tab_table <- lod_tab_summary |>
  pivot_wider(names_from = Value, values_from = Count, values_fill = list(Count = 0)) |> 
  filter(Success > 80) |> 
  select(Element, Success, "< LOD")
lod_tab_table_long <- lod_tab_table |> 
  pivot_longer(cols = c(Success, `< LOD`), names_to = "Metric", values_to = "Count") |> 
  mutate(Metric = factor(Metric, levels = c("Success", "< LOD"))) 
lod_tab_table_long |> 
  ggplot(aes(x = Element, y = Count, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Success and < LOD", y = "Count") +
  theme_minimal() +
  scale_fill_manual(values = c("Success" = "#99D3A7", "< LOD" = "#ED7673"))

# Create the bar plot
ggplot(lod_tab_summary, aes(x = Element, y = Count, fill = Value)) +
  geom_bar(stat = "identity", position = "stack") +  # Use stat = "identity" to use summarized counts
  scale_fill_manual(values = c("Success" = "darkseagreen", "Success 0" = "lightgoldenrod1", "< LOD" = "indianred3")) +
  labs(title = "Summarized Success and LOD Counts by Element",
       y = "Count", 
       fill = " ") +
  theme_minimal() +
 # facet_wrap(~ Element, scales = "free_y") +  # Facet by Element with independent y-axes
  theme(axis.text.x = element_text(hjust = 1),
        axis.title.x = element_blank()) +
  #geom_text(data = lod_tab_summary_filtered, aes(label = Count), position=position_stack(vjust=0.5), colour="white", size = 3) +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, max(lod_tab_summary$Count), by = 20))

```

```{r}
#| label: pxrf-PCA
#| include: false

## a) PCA geological regional cherts
pca_db_algarve <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, S, Cl, K2O, CaO, TiO2, Fe2O3) |>
  filter(`Formation/Type` %in% c("LJW", "MJE", "UJE", "UJW", "TW"))
# Convert "< LOD" values to NA
pca_db_algarve[pca_db_algarve == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_algarve$`Formation/Type` <- as.factor(pca_db_algarve$`Formation/Type`)  # Ensure it's a factor
group_var_algarve <- pca_db_algarve$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_alg <- pca_db_algarve %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_alg <- complete.cases(df_numeric_alg)  # Identify complete rows
df_numeric_alg <- df_numeric_alg[complete_cases_alg, ]  # Remove NA rows
group_var_algarve <- group_var_algarve[complete_cases_alg]  # Filter group_var accordingly
# Scale values
df_scaled_alg <- scale(df_numeric_alg)
# Ensure row names are properly set
rownames(df_scaled_alg) <- make.names(group_var_algarve, unique = TRUE)
# Perform PCA
pca_result_alg <- PCA(df_scaled_alg, graph = FALSE)
# Summary of PCA
summary(pca_result_alg)
fviz_eig(pca_result_alg, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_alg <- fviz_pca_biplot(pca_result_alg, 
                label = "var", 
                habillage = group_var_algarve,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.95) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_alg <- data.frame(pca_result_alg$ind$coord, Group = group_var_algarve)
# PCA plot with groups
scatter_alg <- ggplot(df_pca_alg, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## b) PCA of geological cherts: local Lower Jurassic, Central Portugal and Cadiz
pca_db_nl <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  filter(`Formation/Type` %in% c("LJW", "CPT", "UCA"))
# Convert "< LOD" values to NA
pca_db_nl[pca_db_nl == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_nl$`Formation/Type` <- as.factor(pca_db_nl$`Formation/Type`)  # Ensure it's a factor
group_var_nl <- pca_db_nl$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_nl <- pca_db_nl %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_nl <- complete.cases(df_numeric_nl)  # Identify complete rows
df_numeric_nl <- df_numeric_nl[complete_cases_nl, ]  # Remove NA rows
group_var_nl <- group_var_nl[complete_cases_nl]  # Filter group_var accordingly
# Scale values
df_scaled_nl <- scale(df_numeric_nl)
# Ensure row names are properly set
rownames(df_scaled_nl) <- make.names(group_var_nl, unique = TRUE)
# Perform PCA
pca_result_nl <- PCA(df_scaled_nl, graph = FALSE)
# Summary of PCA
summary(pca_result_nl)
fviz_eig(pca_result_nl, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_nl <- fviz_pca_biplot(pca_result_nl, 
                label = "var", 
                habillage = group_var_nl,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.95) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_nl <- data.frame(pca_result_nl$ind$coord, Group = group_var_nl)
# PCA plot with groups
scatter_nl <- ggplot(df_pca_nl, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## c) PCA archaeological T6 compared to local Lower Jurassic and Central Portugal
pca_db_comp <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2",
    `Formation/Type` == "TYPE 3" ~ "T3",
    `Formation/Type` == "TYPE 4" ~ "T4",
    `Formation/Type` == "TYPE 5" ~ "T5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  )) |> 
  filter(`Formation/Type` %in% c("LJW","CPT", "T6"))
# Convert "< LOD" values to NA
pca_db_comp[pca_db_comp == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_comp$`Formation/Type` <- as.factor(pca_db_comp$`Formation/Type`)  # Ensure it's a factor
group_var_comp <- pca_db_comp$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_comp <- pca_db_comp %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_comp <- complete.cases(df_numeric_comp)  # Identify complete rows
df_numeric_comp <- df_numeric_comp[complete_cases_comp, ]  # Remove NA rows
group_var_comp <- group_var_comp[complete_cases_comp]  # Filter group_var accordingly
# Scale values
df_scaled_comp <- scale(df_numeric_comp)
# Ensure row names are properly set
rownames(df_scaled_comp) <- make.names(group_var_comp, unique = TRUE)
# Perform PCA
pca_result_comp <- PCA(df_scaled_comp, graph = FALSE)
# Summary of PCA
summary(pca_result_comp)
fviz_eig(pca_result_comp, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_comp <- fviz_pca_biplot(pca_result_comp, 
                label = "var", 
                habillage = group_var_comp,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.80) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_comp <- data.frame(pca_result_comp$ind$coord, Group = group_var_comp)
# PCA plot with groups
scatter_comp <- ggplot(df_pca_comp, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## d) PCA archaeological local cherts T2-5 and non-local T7 compared with local Lower Jurassic and UCA cherts
pca_db_comp2 <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, TiO2, K2O,Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2-5",
    `Formation/Type` == "TYPE 3" ~ "T2-5",
    `Formation/Type` == "TYPE 4" ~ "T2-5",
    `Formation/Type` == "TYPE 5" ~ "T2-5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  )) |> 
  filter(`Formation/Type` %in% c("LJW","UCA", "T7", "T2-5"))
# Convert "< LOD" values to NA
pca_db_comp2[pca_db_comp2 == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_comp2$`Formation/Type` <- as.factor(pca_db_comp2$`Formation/Type`)  # Ensure it's a factor
group_var_comp2 <- pca_db_comp2$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_comp2 <- pca_db_comp2 %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_comp2 <- complete.cases(df_numeric_comp2)  # Identify complete rows
df_numeric_comp2 <- df_numeric_comp2[complete_cases_comp2, ]  # Remove NA rows
group_var_comp2 <- group_var_comp2[complete_cases_comp2]  # Filter group_var accordingly
# Scale values
df_scaled_comp2 <- scale(df_numeric_comp2)
# Ensure row names are properly set
rownames(df_scaled_comp2) <- make.names(group_var_comp2, unique = TRUE)
# Perform PCA
pca_result_comp2 <- PCA(df_scaled_comp2, graph = FALSE)
# Summary of PCA
summary(pca_result_comp2)
fviz_eig(pca_result_comp2, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_comp2 <- fviz_pca_biplot(pca_result_comp2, 
                label = "var", 
                habillage = group_var_comp2,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.80) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_comp2 <- data.frame(pca_result_comp2$ind$coord, Group = group_var_comp2)
# PCA plot with groups
scatter_comp2 <- ggplot(df_pca_comp2, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(x = "PC1", y = "PC2") +
  theme(legend.position = "right")

```

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

All analysed samples are mainly composed of quartz; in all samples, the atomic percentage of oxygen is approximately double of the atomic percentage of silica, showing a 2:1 ratio (table +).

Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, shows no patterns and occurs in samples from all formations and chert types. Although frequently without a discernible shape, oval fossils (although currently unidentified) were observed in LJW geological samples and archaeological samples from local cherts (T2 and T4). This type of large, oval fossil, also composed of calcium, phosphorus and sulphur, seems to be limited to local cherts and to specific areas of nodules, characterised by high amounts of these skeletal remains.

```{r}
#| label: fig-skeletal
#| fig-cap: "A) Skeletal remains from sample SP16_BLX, from the Lower Jurassic outcrops of western Algarve. B) Detail of skeletal remains composed of calcium (B1), phosphorus (B2) and a small peak of sulphur (B3). The skeletal remains seem to be oxidized, with small peaks of iron within (B4)."

knitr::include_graphics(here::here("figures/within-chert/M3_3_map1.png"))

```

```{r}
#| label: fig-oval-fossils
#| fig-cap: "A-C) General view and details from an archaeological sample from local chert type (T4). A) general view of a concentration of oval fossils. B) general view of a concentration of oval fossils and other inclusions. C) detail of an irregular inclusion composed of titanium (C1) and an oval skeletal remains composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur."

knitr::include_graphics(here::here("figures/within-chert/M15_4.png"))

```

By observing the surfaces of samples and the presence of skeletal remains, some differences could be noted between local cherts and non-local archaeological samples. In comparison to the frequent to common skeletal remains in the Jurassic cherts from the Algarve, and archaeological cherts attributed to local sources (T2-5), cherts from non-local sources such as T6-8 show little porosity and rare presence of inclusions. Whenever present, skeletal remains are small and show small peaks (\<3 cpv/ev).

```{r}
#| label: fig-t6-general
#| fig-cap: "A-D) General view and details from an archaeological sample from non-local chert type (T6). Dots represent the measured spectrum and shown in the corresponding graph. A) general view of the sample with uncommon inclusions. B) detail of an inclusion seen in A, composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil (1). C) general view of the sample with no visible inclusions. D) detail of a small and irregular iron inclusion (2)."

knitr::include_graphics(here::here("figures/within-chert/M18_7_M18_8_graphs.png"))

```

The rarity of inclusions is not limited to skeletal remains, but to other inclusions such as oxides, which in T6-8 are small and rare. Despite this difference in abundance and size, oxides are mainly composed of iron, often associated with other elements, possibly suggesting the existence of iron oxide (Fe2O3), pyrite (FeS2) and ilmenite (FeTiO3). Iron oxides were found both as inclusions scattered in the samples, but also frequently associated with skeletal remains, which may be related to their oxidization. Pyrite was also identified in the shape of spherical pyrite framboids, as well as dispersed pyrite nanocrystals. These are characterised not only by large peaks of iron and associated sulphur, but also of round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids. These were present mainly in geological samples from the Upper Jurassic (both west and eastern formations), but barely identified in archaeological samples.

```{r}
#| label: fig-framboids
#| fig-cap: "A-B) Detail from the geological sample SP59_JOR from the Upper Jurassic outcrops (Jordana formation) of eastern Algarve. A) detailed view of a spherical pyrite framboid composed of iron (A1) and sulphur (A2). B) detailed view of an inclusion composed of iron (B1) and sulphur (B2)."

knitr::include_graphics(here::here("figures/within-chert/M10_5_map1_map2.png"))

```

In a single sample from Lower Jurassic and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO2). Small peaks of titanium (\<3 cpv/ev) were also identified in 3 out of 4 non-local T6 samples.

Other elements were identified through EDS maps and spectra (e.g., chlorine, potassium or magnesium), although their distribution also showed no patterns between formations and chert types. The exception are calcium and aluminium.

Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations and is probably related to the presence of calcite inclusions.

```{r}
#| label: fig-calcite
#| fig-cap: "General view of calcite inclusions (light grey) from the geological sample SP69_MAR from the Upper Jurassic outcrops of western Algarve, composed mainly of calcium and silica (1). Dot represents the measured spectrum and is shown in the corresponding graph."

knitr::include_graphics(here::here("figures/within-chert/M4_3_graphs.png"))

```

Aluminium may be related to the speed of sedimentation; high peaks of aluminum may be related to fastest sedimentations speeds and may be a useful component to identify differences between cherts. Almost all chert samples show aluminum peaks, although mostly small (\< 3 cpv/ev). An exception are cherts from the Eastern basin of the Algarve (MJE and UJE), samples from non-local archaeological T7 and T8. The exception to the presence of aluminum is the chalcedony (both geological from TW and archaeological group T1) and all archaeological samples from the non-local group T6, with sources previously attributed to the Cretaceous cherts from Central Portugal.

```{r}
#| label: fig-pca
#| fig-cap: Several pcas

plot_grid(biplot_alg, biplot_nl, biplot_comp, biplot_comp2, labels = c("a", "b", "c", "d"))

```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```